\vspace{-2cm} \LARGE Divergence-Free Smoothed Particle Hydrodynamics \\ \large Advanced Simulation and Visualization of Fluids in Computer Graphics

### 2016-12-14 report/paper.tex

\subsection{Divergence solver} \label{divergenceSolver}
\subsection{Density solver} \label{densitySolver}
    This solver aims to minimize the density error caused by the numerical integration in the divergence solver. The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$. There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et. al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver. The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}. For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.

@@

% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
The quicker the particles move the smaller time steps are taken such that the numerical 
error of the explicit time integration is negligible. For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.

@@

\subsection{Density and alpha factors}
    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}. This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.

>>>>

\subsection{Divergence solver} \label{divergenceSolver}

\iffalse
\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
function correctDivergenceError
    for all particles $i$ do
        calculateDensityChange $\frac{Dp}{Dt}_i$
    for all particles $i$ do //Scale $\color{codegreen}a_i$
        $a_i = \frac{a_i}{\Delta t}$
    while (t < $t_{max}$) do //Start simulation loop
        adapt time step $\Delta{t}$ 
        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
        for all particles $i$ do //Update positions
            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
        for all particles $i$ do //Update neighbourhoods
            find neighbourhoods $N_i(t + \Delta{t})$
        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
            compute densities $\rho_i(0)$
            compute factors $\alpha_i(0)$
        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
\end{lstlisting}
\fi

The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation %\ref{eq:navier}.

The solver will ensure that the divergence error is of an acceptable level defined by the user. The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$. The pressure force is determined by

\begin{equation} \label{eq:pressure}
\vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i},
\end{equation}

where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}.

\begin{equation} \label{eq:dpdt}
\frac{D\rho_i}{Dt} = \sum_j m_j (\vect{v_i} - \vect{v_j})\nabla \vect{W_{ij}}
\end{equation}

\begin{equation} \label{eq:kappa}
\nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
\end{equation}

Equation \ref{eq:dpdt} defines how the density change is computed and \ref{eq:kappa} computes the pressure gradient with the help of the stiffness parameter $\kappa_i^v$, where

\begin{equation}
\kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
\end{equation}

and $\alpha_i$ is the scaling factor only dependant on the position explained in section \ref{section:alpha}. The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
The velocity of particle $i$ is thus changed to:

\begin{equation} \label{eq:pressuretotal}
\vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} \cdot \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
\end{equation}

After the symmetric pressure force has been applied to the new velocity should the density change in equation \ref{eq:dpdt} be equal or close to zero. If the error is too large will the process be iterated until desired result have been achieved.

\subsection{Density solver} \label{densitySolver}
    This solver aims to minimize the density error caused by the numerical integration in the divergence solver. The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$. There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et. al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver. The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}. For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.

@@

% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
The quicker the particles move the smaller time steps are taken such that the numerical 
error of the explicit time integration is negligible. For this is a condition proposed by 
Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. 
The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
the maximum velocity $\vect{V_{max}}$ in the scene.

@@

\subsection{Density and alpha factors} \label{section:alpha}
    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}. This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
